### 2023-08-30 proofs/tex/singularity-lemmas.tex
WIP: add completeness of singularity

\begin{lemma}[Completeness of Singularity]
\end{lemma}

\begin{lemma}[Completeness of Singularity]
    \label{lemma:singularity-completeness}
    Suppose there exists $[[Θ ⊢ uσ1]]$ such that
    for any $[[Θ ⊢ uσ]]$, $[[Θ ⊢ uσ : SC]]$ implies $[[Θ ⊢ uσ ≈ uσ1 : varset]]$.
    Then 
    \begin{itemize}
        \item $[[SC|varset singular with uσ0]]$ for some $[[uσ0]]$, and
        \item $[[varset ⊆ dom(SC)]]$.
    \end{itemize} 
\end{lemma}
\begin{proof}
\end{proof}

>>>>

\begin{lemma}[Completeness of Singularity]
    \label{lemma:singularity-completeness}
    For a given $[[Θ ⊢ SC]]$,
    suppose that all the substitutions satisfying $[[SC]]$ are equivalent
    on $[[Ξ]] \supseteq [[dom(SC)]]$.
    In other words, suppose that there exists $[[Θ ⊢ uσ1 : Ξ]]$ such that
    for any $[[Θ ⊢ uσ : Ξ]]$, $[[Θ ⊢ uσ : SC]]$ implies 
    $[[Θ ⊢ uσ ≈ uσ1 : Ξ]]$.
    Then 
    \begin{itemize}
        \item $[[SC singular with uσ0]]$ for some $[[uσ0]]$ and
        \item $[[Ξ]] = [[dom(SC)]]$.
    \end{itemize} 
\end{lemma}
\begin{proof}

    First, let us assume $[[Ξ]] \neq [[dom(SC)]]$. 
    Then there exists $[[α̂± ∊ Ξ \ dom(SC)]]$.
    Let us take $[[Θ ⊢ uσ1 : Ξ]]$ such that 
    any other substitution $[[Θ ⊢ uσ : Ξ]]$ satisfying $[[SC]]$
    is equivalent to $[[uσ1]]$ on $[[Ξ]]$.

    Notice that $[[Θ ⊢ uσ1 : SC]]$: 
    by \cref{lemma:constraint-sat}, there exists $[[uσ']]$
    such that $[[Θ ⊢ uσ' : Ξ]]$ and $[[Θ ⊢ uσ' : SC]]$,
    and by assumption, $[[Θ ⊢ uσ' ≈ uσ1 : Ξ]]$,
    implying  $[[Θ ⊢ uσ' ≈ uσ1 : dom(SC)]]$.

    Let us construct $[[uσ2]]$ such that $[[Θ ⊢ uσ2 : Ξ]]$ as
    follows:
    $$
    \begin{cases}
        [[ [uσ2]β̂± = [uσ1]β̂±  ]] & \text{if } [[β̂± ≠ α̂±]] \\
        [[ [uσ2]α̂± ]] = T & \text{where $T$ is any closed type not equivalent to $[[ [uσ1]α̂± ]]$} \\
    \end{cases}
    $$
    It is easy to see that $[[Θ ⊢ uσ2 : SC]]$ since
    $[[uσ1|dom(SC) = uσ2|dom(SC)]]$, and $[[Θ ⊢ uσ1 : SC]]$.
    However, $[[Θ ⊢ uσ2 ≈ uσ1 : Ξ]]$ does not hold because
    by construction, $[[Θ(α̂±) ⊢ [uσ2]α̂± ≈ [uσ1]α̂±]]$ does not hold. 
    This way, we have a contradiction.

    Second, let us show $[[SC singular with uσ0]]$.
    Let us take arbitrary $[[scE ∊ SC]]$ restricting $[[β̂±]]$.
    We need to show that $[[scE]]$ is singular.
    Notice that $[[Θ ⊢ uσ1 : SC]]$ implies $[[Θ(β̂±) ⊢ [uσ1]β̂±]]$ and $[[Θ(β̂±) ⊢ [uσ1]β̂± : scE ]]$.
    We will show that any other type satisfying $[[scE]]$ is equivalent to $[[ [uσ1]β̂± ]]$,
    then by \cref{lemma:entry-singularity-completeness}, $[[scE singular with [uσ1]β̂± ]]$.
    \begin{itemize}
        \item if $[[β̂±]]$ is positive, 
            let us take any type $[[Θ(β̂±) ⊢ iP']]$
            and assume $[[Θ(β̂±) ⊢ iP' : scE]]$.
            We will show that $[[Θ(β̂±) ⊢ iP' ≈ [uσ1]β̂±]]$,
            which by \cref{lemma:subt-equiv-algorithmization} will
            imply $[[scE singular with nf([uσ1]β̂±)]]$.

            Let us construct $[[uσ2]]$ such that $[[Θ ⊢ uσ2 : Ξ]]$ as
            follows:
            $$
            \begin{cases}
                [[ [uσ2]γ̂± = [uσ1]γ̂±  ]] & \text{if } [[γ̂± ≠ β̂±]] \\
                [[ [uσ2]β̂± = iP' ]]
            \end{cases}
            $$
            It is easy to see that $[[Θ ⊢ uσ2 : SC]]$:
            for $[[scE]]$, $[[Θ(β̂±) ⊢ [uσ2]β̂± : scE ]]$ by construction,
            since $[[Θ(β̂±) ⊢ iP' : scE]]$; for any other $[[scE' ∊ SC]]$
            restricting $[[γ̂±]]$, $[[ [uσ2]γ̂± = [uσ1]γ̂± ]]$, 
            and $[[Θ(γ̂±) ⊢ [uσ1]γ̂± : scE' ]]$ since $[[Θ ⊢ uσ1 : SC]]$.

            Then by assumption, $[[Θ ⊢ uσ2 ≈ uσ1 : Ξ]]$,
            which in particular means $[[Θ(β̂±) ⊢ [uσ2]β̂± ≈ [uσ1]β̂±]]$,
            that is $[[Θ(β̂±) ⊢ iP' ≈ [uσ1]β̂±]]$.
        \item if $[[β̂±]]$ is negative, the proof is analogous.
    \end{itemize}
\end{proof}
